---
title: "R Notebook"
output: html_notebook
---

```{r}
setwd("~/Projects/BayesianStatistics/Chapter_17_HierarchicalModels")
```

```{r}
library(rstan)
options(mc.cores = parallel::detectCores())
```

# 17.1.1

```{r}
study <- read.csv('hierarchical_betaBlocker.csv')
study
```

```{r}
dataList <- list(N=22, rt=study$rt, nt=study$nt, rc=study$rc, nc=study$nc, N=study$N)
```

```{r}
fit <- stan(file = 'BetaBlocker.stan', data=dataList, iter=1000, chain=4, seed=1)
```

```{r}
print(fit)
```

#17.1.2. 
An alternative framework is a hierarchical model where we assume there to be a common over-arching distribution, across trials such that δi ∼ N (d, σ). By assuming the following priors on these parameters estimate this model:
d ∼ N (0, 10) (17.9)
σ ∼ Cauchy(0, 2.5), for σ ≥ 0 (17.10) 

Estimate the posteriors for δi using Stan. How do these estimates compare to the non-hierarchical model?

```{r}
fit_hier <- stan(file = 'BetaBlockerHierarchical.stan', data=dataList, iter=1000, chain=4, seed=1,control = list(adapt_delta = 0.95))
```
```{r}
print(fit_hier)
```

# 17.1.3. 
Using the hierarchical model estimate the cross-study effect of the beta-blockers. (Hint: use the generated quantities code block.)
```{r}
params_hier <- extract(fit_hier)
```

```{r}
hist(params_hier$delta_new)
abline(v=0, col="blue")
```


