---
title: "Problem 15.1"
output:
  pdf_document:
    latex_engine: xelatex
  html_document:
    df_print: paged
---

# 15.1.1
Use your statistical software of choice generate 100 independent samples of (μt,μc). Draw a scatter plot of your (μt,μc) samples, with lines connecting consecutive points. How close are the sample-estimated means to the true means? (Hint: to do this in R you will need to use the MASS package:

```{r}
library (MASS)
Sigma <- matrix(c(2,0.8,0.8,0.5),2,2)
hist(mvrnorm(n = 100, c(20,5), Sigma), main = "Mean of mu_t and mu_c")
```

# 15.1.2
Code up a Random Walk Metropolis sampler for this example. This is composed of the following steps:

```{r}
ldebug <- 0
Sigma <- matrix(c(2, 0.8, 0.8, 0.5), 2, 2)
```

## 1 
Create a proposal function that takes a current value of θ = (μt, μc) and outputs a proposed value of these using a multivariate normal centred on the current estimates. (Here use a multivariate normal proposal with an identity covariance matrix.)

```{r}
fproposal <- function(theta_current){
  if (ldebug == 1) cat("\nfproposal:theta_current = ", theta_current)
  return(mvrnorm(n=1, mu = theta_current,Sigma = Sigma))
}
```


## 2
Create a function which takes as inputs θcurrent and θproposed, and outputs the ratio of the posteriors of the proposed value to the current one (Hint: to do this in R you will need to use the following to calculate the value of the posterior at (x, y):

```{r}
library (mvtnorm)
fGetRatio <- function(theta_current, theta_proposed) {
  r <- dmvnorm(theta_proposed, c(20, 5), sigma = Sigma) / dmvnorm(theta_current ,c(20, 5), sigma = Sigma)
  
  if (ldebug == 1) cat("\nfGetRatio:r= ", r)
  
  return(r)

}

```

## 3
Create an accept/reject function which takes as inputs θcurrent and θproposed, then uses the
above ratio function to find: r = θproposed ; then compares r with a uniformly-distributed θcurrent random number u between 0 and 1. If r > u =⇒ output θproposed; otherwise output θcurrent

```{r}
fAcceptReject <- function(theta_current, theta_proposed) {
  r <- fGetRatio(theta_current, theta_proposed)
  
  u <- runif(1,0,1)
  if ( r > u)
   return(theta_proposed)
  else
   return(theta_current)
}

```

## 4
Combine the proposal function along with the accept/reject function to make a function that takes as input θcurrent, proposes a new value of θ, then based on r moves to that new point or stays in the current position.

```{r}

fSample <- function(theta_current) {
  if (ldebug == 1) cat("\nfSample: theta_current= ", theta_current)
  
  theta_proposed <- fproposal(theta_current)
  if (ldebug == 1) cat("\nfSample:theta_proposed= ", theta_proposed)
  
  theta_accepted <- fAcceptReject(theta_current, theta_proposed)
  if (ldebug == 1) cat("\nfSample:theta_accepted= ", theta_accepted)
  
  return(theta_accepted)
  
}

```

## 5
Create a function called “RWMetropolis” that takes a starting value of θ and runs for n steps

```{r}
RWMetropolis <- function(theta, N) {
  if (ldebug == 1) cat("\nRWMetropolis: theta = ", theta)
  
  nRejected <- 0
  nextSample <- matrix(nrow = 1, ncol = 2)
  
  ltheta <- matrix(nrow = N, ncol = 2)
  ltheta[1,] <- theta
  
  for (t in 2:N){
    nextSample[1,] <- fSample(ltheta[t-1,])
  
    ifelse(nextSample[1,] == ltheta[t-1,], nRejected <- nRejected + 1, 0)
    
    ltheta[t,] <- nextSample[1,]
    
  }
  
  result <- list("nRejected" = nRejected , "samples" = ltheta)
  return(result)
}
```

# 15.1.2
```{r}
theta_starting <- c(10,5)
ldebug <- 0
result <- RWMetropolis(theta_starting,100)
noOfRejectedSamples <- result$"nRejected"
lRWMsamples <- result$samples

cat("\n No. of Rejected Samples: ", noOfRejectedSamples)
cat("\n % of rejected samples = ", noOfRejectedSamples/(100 + noOfRejectedSamples) )

```


```{r}
plot(lRWMsamples, type = 'l', xlab = "mu_t", ylab = "mu_c", main = "Random Wlak Metropolis")
```

```{r}
hist(lRWMsamples[,2], col ="blue", xlab = "mu_c")
hist(lRWMsamples[,1], col ="orange", xlab = "mu_t")
```

# 15.1.3
```{r}
acceptancePerc <- vector()
for(i in 1:100) {
  result <- RWMetropolis(theta_starting,100)
  acceptancePerc[i] <- result$"nRejected" / (100 + result$"nRejected")
}

hist(acceptancePerc, xlab = "rejection %", main = "RWM ARejection %", col = "blue")

```

# 15.1.4
```{r}
fWithin <- function(lSamples){
  return(mean(sapply(lSamples, var)))
}
## Testing it
#lSamples <- lapply(seq(1, 10, 1), function(i) rbinom(100, 100, 0.5))
#fWithin(lSamples)
```

```{r}
fBetween <- function(lSamples){
  lMean <- sapply(lSamples, mean)
  aMean <- mean(lMean)
  m <- length(lSamples)
  t <- length(lSamples[[1]])
  return((t / (m - 1)) * sum((lMean - aMean) ^ 2))
}
  
```

```{r}
fRhat <- function(lSamples){
  W <- fWithin(lSamples)
  B <- fBetween(lSamples)
  t <- length(lSamples[[1]])
  return(sqrt((W + (1 / t) * (B - W)) / W))
}
```


# 15.1.5
```{r}
theta_starting <- c(20,5)
numSamples <- 5

acceptancePerc <- vector()

lSamples <- list()

for(i in 1:8) {
  result <- RWMetropolis(theta_starting,numSamples)
  lRWMsamples <- result$samples
  lSamples[[i]] <- lRWMsamples
  acceptancePerc[i] <- result$"nRejected" / (numSamples + result$"nRejected")
}

lRhat <- fRhat(lSamples)
```

```{r}
cat("\nRhat : ", lRhat)
cat("\n Rejected % : ", acceptancePerc)
```

# 15.1.6

```{r}

numSamples <- 1000
acceptancePerc <- vector()
lSamples <- list()

for(i in 1:8) {
  theta_starting <- mvrnorm(n=1, c(20, 5), Sigma = 40 * diag(2))
  result <- RWMetropolis(theta_starting,numSamples)
  lRWMsamples <- result$samples
  lSamples[[i]] <- lRWMsamples
  acceptancePerc[i] <- result$"nRejected" / (numSamples + result$"nRejected")
}

lRhat <- fRhat(lSamples)

```


```{r}
cat("\nRhat : ", lRhat)
cat("\nRejected % : ", acceptancePerc)

```

```{r}
dim(lSamples[[1]])
```


```{r}

mu_t <- matrix(nrow = 8, ncol = numSamples)
mu_c <- matrix(nrow = 8, ncol = numSamples)

for (i in 1:8) {
  mu_t[i,] <- lSamples[[i]][,1]
  mu_c[i,] <- lSamples[[i]][,2]
}


```


```{r}
palette()
```


```{r}

plot(mu_t[1,], type="l", col=1, ylim = c(0, 35), xlab = "iteration#", 
     ylab = expression(paste(mu[t])), 
     main = expression(paste("Figure 15.5: The path taken by each chain in ", μ[t]," space")))

for(i in 2:8){
  lines(mu_t[i,], type="l", col=i)
}

```



